### 2020-05-30 writeup/setup.tex
sec 4

\subsection{A Key Lemma using Random Matrix Theory}\label{label_rmt}
%For the high-dimensional setting where the data size of the target task is a small constant times dimension $p$, we quantify the trade-off as a function of the task models $\set{\beta_i}_{i=1}^t$, data size $\set{n_i}_{i=1}^t$, and covariance matrices $\set{\Sigma_i}_{i=1}^t$.

@@

\begin{theorem}[Informal]\label{thm_main_informal}
	Let $X_i \in\real^{n_i \times p}$ and $Y_i = X_i\beta_i + \varepsilon_i$, for $i = 1, 2$.
	Suppose that $n_1 = \rho_1 p$ and $n_2 = \rho_2 p$, where $\rho_1>1$ and $\rho_2 >1$ are fixed constants.
	There exists two deterministic functions $\Delta_{\beta}$ and $\Delta_{\vari}$ and a small deterministic error $\delta$ that only depend on $\set{\hat{v}, \Sigma_1^{}, n_1, n_2, \beta_1, \beta_2}$ such that
	\begin{itemize}
		\item If $\Delta_{\vari} - \Delta_{\beta} \ge \delta$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Delta_{\vari} - \Delta_{\beta} \le \delta$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{theorem}

Theorem \ref{thm_main_informal} shows nearly tight bounds on the trade-off between model-shift bias and variance reduction.
%determined by the covariate shift matrix and the model shift.
The bounds get tighter and tighter as $\rho_1$ increases.
While the general form of $\Delta_{\vari}$ and $\Delta_{\beta}$ can be quite complex, we will show that they provide nice interpretation for simplified settings later on in Section \ref{sec_insight}.

We describe an overview of the proof of Theorem \ref{thm_main_informal}.
By comparing the bias and variance of $\te_t(\hat{\beta}_t^{\MTL})$ to $\te_t(\hat{\beta}_t^{\STL})$, we observe the following two effects, respectively.
%We begin by observing that the test error of $\hat{\beta}_t^{\MTL}$ consists of two parts.
We term equation \eqref{eq_te_model_shift} as \textit{model-shift bias}, which captures how similar $\beta_1$ and $\beta_2$ are.
This part introduces a negative effect to multi-task learning.
The scaling term $\hat{v}$ corresponds to the intuition that the shared module $B$ learns a subspace for both tasks.
The output layer of each task scales the direction of $B$ suitably to fit the task.

Next, it is not hard to verify that equation \eqref{eq_te_var}, the variance of $\hat{\beta}_t^{\MTL}$, is always smaller than the variance of $\hat{\beta}_t^{\STL}$.
This part introduces a positive variance reduction effect to performing multi-task learning.
Hence, whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the trade-off between two effects:
(i) the positive effect from variance reduction;
(ii) the negative effect from model shift bias.
A formal version of Theorem \ref{thm_main_informal} is presented in Theorem \ref{thm_model_shift} and its proof is presented in Appendix \ref{app_proof_main}.

>>>>

\subsection{A Key Lemma using Random Matrix Theory}\label{label_rmt}

\begin{lemma}[Informal statement of Lemma \ref{lem_cov_shift}]\label{lem_cov_shift_informal}
	Suppose $X_1=Z_1\Sigma_1^{1/2}\in \R^{n_1\times p}$ and $X_2=Z_2\Sigma_2^{1/2}\in \R^{n_2\times p}$ with $\rho_1=n_1/p>1$ and $\rho_2=n_2/p>1$.
	Let $M = \Sigma_1^{1/2}\Sigma_2^{-1/2}$ and $\lambda_1, \lambda_2, \dots, \lambda_p$ be the singular values of $M^{\top}M$ in descending order.
	For any constant $\e>0$, w.h.p. over the randomness of $X_1, X_2$, we have that
	\begin{align*}
		\bigtr{(X_1^{\top}X_1 + X_2^{\top}X_2)^{-1}\Sigma_2} = \frac{1}{\rho_1+\rho_2}\cdot \frac1p\bigtr{ (a_1 \Sigma_1 + a_2\Sigma_2)^{-1} \Sigma_2} +\bigo{\|\Sigma_2\| p^{-1/2+\epsilon}},
	\end{align*}
where $(a_1, a_2)$ is the solution to the following deterministic equations:
	\begin{align*}
		a_1 + a_2 = 1- \frac{1}{\rho_1 + \rho_2},\quad a_1 + \frac1{\rho_1 + \rho_2}\cdot \frac{1}{p}\sum_{i=1}^p \frac{\lambda_i^2 a_1}{\lambda_i^2 a_1 + a_2} = \frac{\rho_1}{\rho_1 + \rho_2}.
	\end{align*}
\end{lemma}

@@

\begin{theorem}[Informal statement of Theorem \ref{thm_model_shift}]\label{thm_main_informal}
	Let $X_i \in\real^{n_i \times p}$ and $Y_i = X_i\beta_i + \varepsilon_i$, for $i = 1, 2$.
	Suppose that $n_1 = \rho_1 p$ and $n_2 = \rho_2 p$, where $\rho_1>1$ and $\rho_2 >1$ are fixed constants.
	There exists two deterministic functions $\Delta_{\beta}$ and $\Delta_{\vari}$ and a small deterministic error $\delta$ that only depend on $\set{\hat{v}, \Sigma_1^{}, n_1, n_2, \beta_1, \beta_2}$ such that
	\begin{itemize}
		\item If $\Delta_{\vari} - \Delta_{\beta} \ge \delta$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Delta_{\vari} - \Delta_{\beta} \le \delta$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{theorem}

Theorem \ref{thm_main_informal} shows nearly tight bounds on the trade-off between model-shift bias and variance reduction.
%determined by the covariate shift matrix and the model shift.
The bounds get tighter and tighter as $\rho_1$ increases.
While the general form of $\Delta_{\vari}$ and $\Delta_{\beta}$ can be quite complex, we will show that they provide nice interpretation for simplified settings later on in Section \ref{sec_insight}.
A formal version of Theorem \ref{thm_main_informal} is presented in Theorem \ref{thm_model_shift} and its proof is presented in Appendix \ref{app_proof_main}.

### 2020-05-30 writeup/transfer.tex
sec 4

\textbf{Cause 1.}
We measure model dissimilarity as $\norm{\beta_1 - \beta_2}^2$, which is the distance between source and target in the isotropic model.
We derive a sharp threshold when positive transfer transitions to negative transfer, as model dissimilarity increases.
%Based on Theorem \ref{thm_model_shift}, we derive the transition threshold in the following proposition.
Define the following function
\begin{align*}
	\Phi(\rho_1, \rho_2) = \frac{(\rho_1 + \rho_2 - 1)^2}{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}.
\end{align*}

\begin{proposition}[Model dissimilarity]\label{prop_dist_transition}
	In the isotropic model, assume that $\rho_1 > 40$ is a fixed constant.
	Let $\sigma_1 = \sigma_2 = \sigma$.
	Whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the ratio between model dissimilarity and $\Phi(\rho_1, \rho_2)$:
	\begin{itemize}
		\item If ${p d^2} < \frac{1}{2}  {\sigma^2}  \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If ${p d^2} \ge 2 {\sigma^2} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}

@@

@@

\begin{proposition}[Source data size]\label{prop_data_size}
	In the isotropic model, assume that $\rho_1 > 40$ and $\rho_2 > 500$ are fixed constants.
	We have the following conditions to determine whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$:
	\begin{itemize}
		\item[a)] If $pd^2 \le \frac{1}{2}\cdot \frac{\sigma^2}{\rho_2-1}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$ for any $\rho_1 > 40$.
		\item[b)] If $pd^2 > 2\cdot\frac{\sigma^2}{\rho_2 - 1}$, then we have the following transition depending on $\rho_1$:
		\begin{itemize}
			\item If $\rho_1 \cdot p < \frac{(\rho_2-2)\sigma^2}{(1 + {\rho_1}^{-0.5})^4(\rho_2 - 1) d^2 - \sigma^2 / p}$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
			\item If $\rho_1 \cdot p > \frac{(\rho_2-2) \sigma^2}{(1 - {\rho_1}^{-0.5})^4 (\rho_2 - 3) d^2 - \sigma^2 / p}$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
		\end{itemize}
	\end{itemize}
\end{proposition}
The intuition behind Proposition \ref{prop_data_size} is as follows.
a) If $\norm{\beta_1 - \beta_2}^2 \approx pd^2$ is sufficiently small, then adding any amount of labeled data from the source task always provides positive transfer;
ii) Otherwise, as source data size increases, we observe a transition from positive to negative transfer.
The proof of Proposition \ref{prop_data_size} is similar to Proposition \ref{prop_dist_transition}, and can be found in Appendix \ref{app_proof_31}.

@@

\begin{proposition}[Labeled data de-noising]\label{prop_var_transition}
	In the isotropic model, assume that $\rho_1 > 40$ and $pd^2 < \frac{1}{2} {\sigma_2^2}  \cdot \Phi(\rho_1, \rho_2)$.
	We derive the following transition as a parameter of $\sigma_1^2$:
	\begin{itemize}
		\item If $\sigma_1^2 \le -2 \rho_1 \cdot p d^2+\left(1+ \frac34\rho_1 \Phi(\rho_1, \rho_2)\right)\cdot\sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\sigma_1^2 > - \frac12\rho_1\cdot p d^2   +\left(1+ \frac32\rho_1\Phi(\rho_1, \rho_2)\right) \cdot \sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}
As a corollary, if $\sigma_1^2 \le \sigma_2^2$, then we always get positive transfer.
The proof of Proposition \ref{prop_var_transition} is similar to Proposition \ref{prop_dist_transition}.
The details can be found in Appendix \ref{app_proof_data}.

>>>>

\textbf{Cause 1.}
We measure model dissimilarity as $\norm{\beta_1 - \beta_2}^2$, which is the distance between source and target in the isotropic model.
We derive a sharp threshold when positive transfer transitions to negative transfer, as model dissimilarity increases.
%Based on Theorem \ref{thm_model_shift}, we derive the transition threshold in the following proposition.
Define the following functions
\begin{align*}
	\Psi(\beta_1, \beta_2) = {\ex{\bignorm{\beta_1 - \beta_2}^2}} / {\sigma^2},  \Phi(\rho_1, \rho_2) = \frac{(\rho_1 + \rho_2 - 1)^2}{\rho_1 (\rho_1 + \rho_2) (\rho_2 - 1)}.
\end{align*}
Let $\alpha_{-} = (1 - \rho_1^{-1/2})^4$ and $\alpha_{+} = (1 + \rho_1^{-1/2})^4$.

\begin{proposition}[Model dissimilarity]\label{prop_dist_transition}
	In the isotropic model, suppose that $\rho_1 > 1$ and $\sigma_1 = \sigma_2 = \sigma$.
	Then we have that
	%Whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the ratio between $\Psi(\beta_1, \beta_2)$ and $\Phi(\rho_1, \rho_2)$:
	\squishlist
		\item If $\Psi(\beta_1, \beta_2) < \alpha_{-} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\Psi(\beta_1, \beta_2) \ge \alpha_{+} \cdot \Phi(\rho_1, \rho_2)$, then whp we have that $\te(\hat{\beta}_t^{\MTL}) \ge \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{proposition}

@@

We describe an overview of the proof of Theorem \ref{thm_main_informal}.
By comparing the bias and variance of $\te_t(\hat{\beta}_t^{\MTL})$ to $\te_t(\hat{\beta}_t^{\STL})$, we observe the following two effects, respectively.
%We begin by observing that the test error of $\hat{\beta}_t^{\MTL}$ consists of two parts.
We term equation \eqref{eq_te_model_shift} as \textit{model-shift bias}, which captures how similar $\beta_1$ and $\beta_2$ are.
This part introduces a negative effect to multi-task learning.
The scaling term $\hat{v}$ corresponds to the intuition that the shared module $B$ learns a subspace for both tasks.
The output layer of each task scales the direction of $B$ suitably to fit the task.
Next, it is not hard to verify that equation \eqref{eq_te_var}, the variance of $\hat{\beta}_t^{\MTL}$, is always smaller than the variance of $\hat{\beta}_t^{\STL}$.
This part introduces a positive variance reduction effect to performing multi-task learning.
Hence, whether $\te(\hat{\beta}_t^{\MTL})$ is lower than $\te(\hat{\beta}_t^{\STL})$ is determined by the trade-off between two effects:
(i) the positive effect from variance reduction;
(ii) the negative effect from model shift bias.

@@

\begin{proposition}[Source data size]\label{prop_data_size}
	In the isotropic model, assume that $\rho_1 > 40$ and $\rho_2 > 500$ are fixed constants, $\Psi(\beta_1, \beta_2) > 2/(\rho_2 - 1)$.
	We have that
	\squishlist
		\item If $\rho_1 < \frac{\rho_2-2}{\alpha_+ \cdot \Phi(\beta_1, \beta_2) (\rho_2 - 1) - 1}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\rho_1 > \frac{\rho_2-2}{\alpha_- \cdot \Phi(\beta_1, \beta_2) (\rho_2 - 3) - 1}$, then w.h.p. $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\squishend
\end{proposition}
The intuition behind Proposition \ref{prop_data_size} is as follows.
a) If $\norm{\beta_1 - \beta_2}^2 \approx pd^2$ is sufficiently small, then adding any amount of labeled data from the source task always provides positive transfer;
ii) Otherwise, as source data size increases, we observe a transition from positive to negative transfer.
The proof of Proposition \ref{prop_data_size} is similar to Proposition \ref{prop_dist_transition}, and can be found in Appendix \ref{app_proof_31}.

@@

\begin{proposition}[Labeled data de-noising]\label{prop_var_transition}
	In the isotropic model, assume that $\rho_1 > 40$ and $\ex{\norm{\beta_1 - \beta_2}^2}< \frac{1}{2} {\sigma_2^2}  \cdot \Phi(\rho_1, \rho_2)$.
	We derive the following transition as a parameter of $\sigma_1^2$:
	\begin{itemize}
		\item If $\sigma_1^2 \le -2 \rho_1 \cdot p d^2+\left(1+ \frac34\rho_1 \Phi(\rho_1, \rho_2)\right)\cdot\sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) < \te(\hat{\beta}_t^{\STL})$.
		\item If $\sigma_1^2 > - \frac12\rho_1\cdot p d^2   +\left(1+ \frac32\rho_1\Phi(\rho_1, \rho_2)\right) \cdot \sigma_2^2$, then whp $\te(\hat{\beta}_t^{\MTL}) > \te(\hat{\beta}_t^{\STL})$.
	\end{itemize}
\end{proposition}
As a corollary, if $\sigma_1^2 \le \sigma_2^2$, then we always get positive transfer.
The proof of Proposition \ref{prop_var_transition} is similar to Proposition \ref{prop_dist_transition}.
The details can be found in Appendix \ref{app_proof_data}.
